---
title: 'CSCI E-63C: Week 2 Q&A session'
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(ISLR)
library(reshape2)
knitr::opts_chunk$set(echo = TRUE)
```

# Questions

* Week 1 homework
* Week 1 lecture
* Upcoming homework
* Upcoming quiz

# Simple ggplot example

```{r,fig.width=12,fig.height=4}
dfTmp <- data.frame(xvals=rnorm(30),yvals=runif(30),nn=rep(c(5,10,20),10))
ggplot(dfTmp,aes(x=xvals,y=yvals,colour=factor(nn)))+geom_point()+facet_wrap(~nn)
```

```{r,fig.width=12,fig.height=6}
ggplot(dfTmp[dfTmp$nn!=5,],aes(x=xvals,y=yvals,colour=factor(nn)))+geom_point()+facet_wrap(~nn)
```


# Factors and type conversion example

```{r typeconversion}
x <- 1:5; someLetters <- letters[seq_along(x)]; someLetters
names(x) <- someLetters; x
names(x) <- sample(names(x)); x
x[someLetters]
fSomeLetters <- factor(someLetters); fSomeLetters
as.character(fSomeLetters); x[as.character(fSomeLetters)]
x[fSomeLetters]
```

# Some of the R conveniences

* `ggplot2` -- popular plotting package providing unified interface to variety of visualizations (e.g. don't have to remember how to combine `boxplot` and `stripchart`)
* tidy data: `reshape2`, `plyr`, ... -- very convenient together with `ggplot`, standardizes many common data transformations
* OOP: class is just an attribute that is easy to change, s3/s4/rc frameworks, dots in names, http://adv-r.had.co.nz/OO-essentials.html
* `shiny` -- enabling interactive R over web

# S3 OOP in R: a word of caution

```{r s3oop}
fSomeLetters
class(fSomeLetters)
is.object(fSomeLetters) & !isS4(fSomeLetters)
class(fSomeLetters) <- "Let's get wild here"; class(fSomeLetters)
fSomeLetters
class(fSomeLetters) <- c("One class","another class"); fSomeLetters
# attribute "class" determines which "print" is called:
class(fSomeLetters) <- "factor"; fSomeLetters
```



# ISLR Ch.2.3.4 (Lab)

```{r ch234}
##?Auto
class(Auto)
dim(Auto)
summary(Auto)
head(Auto)
names(Auto)
colnames(Auto)
# ISLR says there are 397 rows, five of which include ?'s:
for ( tmpCnm in colnames(Auto) ) {
  cat(tmpCnm,class(Auto[,tmpCnm]),sum(Auto[,tmpCnm]=="?"),fill=TRUE)
}
#?write.table
#?read.table
write.table(Auto,"Auto.txt")
autoDat <- read.table("Auto.txt")
head(autoDat)
dim(autoDat)
dim(na.omit(autoDat))
if ( FALSE ) {
  plot(cylinders, mpg)
}
plot(Auto$cylinders, Auto$mpg)
attach (Auto)
plot(cylinders, mpg)
detach (Auto)
plot(Auto$cylinders, Auto$mpg)
ggplot(Auto,aes(x=cylinders,y=mpg))+geom_point()
ggplot(Auto,aes(x=cylinders,y=mpg))+geom_point()+theme_bw()
plot(factor(Auto$cylinders), Auto$mpg, col ="red", varwidth=T, xlab="cylinders", ylab="MPG")
ggplot(Auto,aes(x=factor(cylinders),y=mpg))+geom_point()+ggtitle("Add plot title here")+xlab("cylinders")+ylab("Miles per gallon, MPG")
ggplot(Auto,aes(x=factor(cylinders),y=mpg))+geom_point()+geom_boxplot()+theme_bw()
ggplot(Auto,aes(x=factor(cylinders),y=mpg))+geom_boxplot()+geom_point()+theme_bw()
hist(Auto$mpg)
hist(Auto$mpg,col="red")
ggplot(Auto,aes(x=mpg))+geom_histogram()+theme_bw()
ggplot(Auto,aes(x=mpg))+geom_histogram(fill="lightgreen",colour="black",bins=30)+theme_bw()
```

```{r,fig.width=12,fig.height=12}
pairs(Auto)
library(GGally)
ggpairs(Auto[,-ncol(Auto)])
summary(Auto)
summary(Auto$mpg)
```

# ISRL Ch.2.4 (Exercises -- applied)

## Problem 8 (a-b)

About reading data from the file and viewing it using `fix` data editor function

## Problem 8 (c)

*i. Use the `summary()` function to produce a numerical summary of the variables in the data set.*

```{r ch24problem8ci}
summary(College)
head(College)
```

*ii. Use the `pairs()` function to produce a scatterplot matrix of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix `A` using `A[,1:10]`.*

```{r ch24problem8cii,fig.width=12,fig.height=12}
pairs(College[,1:10])
ggpairs(College[,1:10])
```

*iii. Use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Private`.*

```{r ch24problem8ciii}
plot(College$Private,College$Outstate)
boxplot(Outstate~Private,College)
class(College$Private)
table(College$Private)
class(College$Outstate)
ggplot(College,aes(x=Private,y=Outstate)) + geom_boxplot()
```

*iv. Create a new qualitative variable, called `Elite`, by binning the `Top10perc` variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school classes exceeds 50%.*

*Use the `summary()` function to see how many elite universities there are. Now use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Elite`.*

```{r}
ggplot(College,aes(x=Top10perc))+geom_histogram(colour="red",fill="white")+theme_bw()
# another way to accomplish what is shown in the textbook:
College$Elite <- factor(c("No","Yes")[as.numeric(College$Top10perc>50)+1])
summary(College$Elite)
table(College$Elite)
old.par <- par(mfrow=c(1,2),ps=16)
plot(College$Elite,College$Outstate,xlab="Elite")
plot(College$Private,College$Outstate,xlab="Private")
par(old.par)
head(melt(College[,c("Elite","Private","Outstate")]))
head(melt(College[,c("Elite","Private","Outstate")],id="Outstate"))
ggplot(melt(College[,c("Elite","Private","Outstate")],id="Outstate"),aes(x=value,y=Outstate))+geom_boxplot()+facet_wrap(~variable)
ggplot(melt(College[,c("Elite","Private","Outstate")]),aes(x=Private,y=value,fill=Elite,colour=Elite))+geom_boxplot(colour="black")+facet_wrap(~variable)+theme_bw()
ggplot(melt(College[,c("Elite","Private","Outstate")]),aes(x=Private,y=value,fill=Elite,colour=Elite))+geom_boxplot(fill="white",outlier.colour=NA)+geom_point(position=position_jitterdodge())+facet_wrap(~variable)
table(College[,c("Elite","Private")])
fisher.test(table(College[,c("Elite","Private")]))
chisq.test(table(College[,c("Elite","Private")]))
summary(lm(Outstate~Private*Elite,College))
```

## Problem 9

**This exercise involves the `Auto` data set studied in the lab. Make sure that the missing values have been removed from the data.**

### Problem 9 (a)

*Which of the predictors are quantitative, and which are qualitative?*

```{r ch24problem9a}
head(Auto)
summary(Auto)
```

### Problem 9 (b)

*What is the range of each quantitative predictor? You can answer this using the `range()` function.*

```{r ch24problem9b}
apply(Auto[,1:7],2,range)
```

### Problem 9 (c)

*What is the mean and standard deviation of each quantitative predictor?*

```{r ch24problem9c}
apply(Auto[,1:7],2,mean)
apply(Auto[,1:7],2,median)
apply(Auto[,1:7],2,sd)
apply(Auto[,1:7],2,mad)
if ( FALSE ) {
  # returns NA's for every column - why?
  apply(Auto,2,mean)
}
```

### Problem 9 (d)

*Now remove the 10th through 85th observations. What is the range, mean, and standard deviation of each predictor in the subset of the data that remains?*

```{r ch24problem9d}
apply(Auto[-(10:85),1:7],2,range)
apply(Auto[-(10:85),1:7],2,mean)
apply(Auto[-(10:85),1:7],2,sd)
```

### Problem 9 (e)

*Using the full data set, investigate the predictors graphically, using scatterplots or other tools of your choice. Create some plots highlighting the relationships among the predictors. Comment on your findings.*

```{r ch24problem9e,fig.width=12,fig.height=12}
pairs(Auto)
ggpairs(Auto[,-ncol(Auto)])
```

### Problem 9 (f)

*Suppose that we wish to predict gas mileage (`mpg`) on the basis of the other variables. Do your plots suggest that any of the other variables might be useful in predicting `mpg`? Justify your answer.*


